% !TEX root = ../thesis.tex

Over the past three months a lot of people have put input into my work. There's two categories of people I need to thank and so I will thank in no particular order. Each of you have made this dissertation what it is and I am so very thankful to each of you. \\

\noindent
Academically. Thankyou to Dr Kaibo Hu, my supervisor, for providing a space for me to project all of my thoughts onto the board and providing useful and thought provoking ideas and comments. Thankyou to Professor Darryl Holm for providing input into a project he wasn't supervising and allowing me to attend his Friday meetings. Further to Dr Oliver Street, Ruiao Hu and all of the STOUP team for their insightful discussions and hospitality when I visited Imperial College. Thanks to Dr Hamid Alemi Ardakani, my future PhD supervisor, for your interesting extensions to my work and further providing a space for me to talk about my thesis. Thanks to Dr Andrew Gilbert, for your interest and also your comments when I was working out how boundary conditions worked. Thanks to Professor Tom Bridges and Dr Cesare Tronci for sparking my joy for the subject when I visited last year. Thanks to all the academics at Oxford this year that I took the courses, especially to the academics whose courses became my special topics, thankyou for reading my reports that can be summarised as `How can you apply manifolds and Geometric Mechanics to this area?'. Finally, thankyou to Kathryn Gillow for providing the course and the space to explore this amazing subject!\\

\noindent
Personally. Thankyou to my parents and Molly who have constantly been reminding me that this dissertation needed to be actually written and wasn't just about my notebook of mathematics. Thanks to my friends here at Oxford and elsewhere. To the people here at Oxford; Usman, Adam, Jad, Nicholas, Peter and Chloe from maths and Christina and Max from Teddies and the Hashers, thanks for dealing with my incessantly talking about second-countable householder spaces. Further thankyou to Matthew for allowing me to rant when my Lagrangians weren't reducing or my Hamiltonians weren't yielding the right brackets and dragging me away from the work when it started to drive me up the wall. Thanks to Colin Wright and the other maths communicators for always being interested in my dissertation and generally interested in what I am doing, I'm looking forward to being back on stage communicating this thesis in the coming months.